Developmental trajectories of word recognition: a priming study in bilinguals

### 2023-11-04 manuscript/manuscript.qmd
Fix Oxford resutls in manuscript

targets::tar_load(
	c(participants,
	  stimuli,
	  vocabulary,
	  attrition_trials,
	  attrition_participants,
	  bvq_data,
	  gaze,
	  data_bcn
	))

targets::tar_load(
	c(model_fits_bcn,
	  model_loos_bcn
	))

@@

```{r dataset}
n_trials <- nrow(attrition)
n_sessions <- n_distinct(attrition$session_id)
n_participants <- n_distinct(attrition$child_id)
n_trials_valid <- nrow(filter(attrition,
							  is_valid_participant,
							  is_valid_trial))
n_sessions_valid <- sum(attrition_participants$is_valid_participant)
n_participants_valid <- n_distinct(
	attrition[attrition$is_valid_participant, ]$child_id)

n_exc_prime <- sum(!attrition$is_valid_gaze_prime)
n_exc_test <- sum(!attrition$is_valid_gaze_test)
n_exc_test_each <- sum(!attrition$is_valid_gaze_test_each)
n_exc_participants <- sum(!attrition_participants$is_valid_participant)

n_longitudinal <- attrition |>
	filter(is_valid_participant,
		   is_valid_trial) |>
	distinct(child_id, session_n) |> 
	count(child_id, name = "times") |> 
	count(times)
```

@@

```{r tbl-attrition}
#| label: tbl-attrition
#| tbl-cap: "Participant-level and trial-level sample size after applying inclusion criteria. The number of excluded participants and trials is indicated between parentheses."
expanded <- expand(attrition, lp, trial_type, is_valid_trial)

n_lp <- attrition |> 
	distinct(is_valid_participant, lp, session_id) |> 
	count(is_valid_participant, lp) |> 
	pivot_wider(names_from = is_valid_participant,
				values_from = n,
				names_repair = janitor::make_clean_names) |> 
	mutate(n_participants_true = true,
		   n_participants_false = false) |> 
	select(lp, true, false)

attrition |> 
	summarise(n_trials_true = sum(is_valid_trial),
			  n_trials_false = sum(!is_valid_trial),
			  .by = c(lp, trial_type)) |> 
	arrange(desc(lp)) |> 
	# across(n_trials, lst(mean, sd, sum), na.rm = TRUE),

	pivot_wider(id_cols = lp,
				names_from = c(trial_type),
				values_from = matches("trials"),
				names_repair = janitor::make_clean_names) |> 
	left_join(n_lp, by = join_by(lp)) |> 
	mutate(lp = case_when(
		lp=="Monolingual" ~ "Monolingual (Catalan or Spanish)",
		lp=="Bilingual" ~ "Bilingual (Catalan-Spanish)",
		.default = lp)) |>
	select(lp, true, false,
		   matches("n_trials")) |> 
	gt(rowname_col = "lp") |> 
	tab_spanner("Participants", matches("participant")) |> 
	tab_spanner("Trials", matches("trial")) |> 
	fmt_integer(everything()) |> 
	cols_merge(c(true, false),
			   pattern = "{1} ({2})") |> 
	cols_merge(c(n_trials_true_cognate,
				 n_trials_false_cognate),
				 pattern = "{1} ({2})") |> 
	cols_merge(c(n_trials_true_non_cognate,
				 n_trials_false_non_cognate),
				 pattern = "{1} ({2})") |> 
	cols_merge(c(n_trials_true_unrelated, 
				 n_trials_false_unrelated),
				 pattern = "{1} ({2})") |> 
	cols_label(true = "N", 
			   n_trials_true_cognate = "Cognate",
			   n_trials_true_non_cognate = "Non-cognate",
			   n_trials_true_unrelated = "Unrelated") |> 
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |>
	grand_summary_rows(fns = list(label = md("*N*")) ~ sum(.),
					   fmt = ~ fmt_integer(.))
```

@@

We collected data from `r n_participants_total$n[1]` children (`r n_participants_sex[1]` female, `r n_participants_sex[2]` male, `r n_participants_sex[3]` sex not reported, age: *Mean* = `r round(ages$age_mean, 2)` months, *SD* = `r round(ages$age_sd, 2)`, *Range* = [`r round(ages$age_min, 2)`-`r round(ages$age_max, 2)`]), living in the Oxfordshire area (United Kingdom), tested at the Oxford BabyLab at the University of Oxford. Families were recruited from maternity rooms in private hospitals and social media, and contacted via phone when the child's age spanned between 20 and 32 months. From the `r n_participants_total$n` children that participated, `r n_participants_sessions$session_1` participated once, and `r n_participants_sessions$session_2` participated twice. Recurrent participants were tested with at least XXX months of difference. We gathered a total of `r n_sessions_total$n` testing sessions. All participants were being raised in exclusively British English monolingual homes.

@@

>>>>

targets::tar_load(
	c(participants,
	  stimuli,
	  vocabulary,
	  attrition_trials,
	  attrition_participants,
	  bvq_data,
	  gaze,
	  data_bcn,
	  data_oxf
	))

targets::tar_load(
	c(model_fits_bcn,
	  model_loos_bcn,
	  model_fits_oxf,
	  model_loos_oxf
	))

@@

```{r dataset}
attrition_bcn <- filter(attrition, location=="Barcelona")
n_trials <- nrow(attrition_bcn)
n_sessions <- n_distinct(attrition_bcn$session_id)
n_participants <- n_distinct(attrition_bcn$child_id)
n_trials_valid <- nrow(filter(attrition_bcn,
							  is_valid_participant,
							  is_valid_trial))
n_sessions_valid <- sum(distinct(attrition_bcn,
								 session_id,
								 .keep_all = TRUE)$is_valid_participant)
n_participants_valid <- n_distinct(
	attrition_bcn[attrition_bcn$is_valid_participant, ]$child_id)

n_exc_prime <- sum(!attrition_bcn$is_valid_gaze_prime)
n_exc_test <- sum(!attrition_bcn$is_valid_gaze_test)
n_exc_test_each <- sum(!attrition_bcn$is_valid_gaze_test_each)
n_exc_participants <- sum(!distinct(attrition_bcn,
									session_id,
									.keep_all = TRUE)$is_valid_participant)

n_longitudinal <- attrition_bcn |>
	filter(is_valid_participant,
		   is_valid_trial) |>
	distinct(child_id, session_n) |> 
	count(child_id, name = "times") |> 
	count(times)
```

@@

```{r tbl-attrition}
#| label: tbl-attrition
#| tbl-cap: "Participant-level and trial-level sample size after applying inclusion criteria. The number of excluded participants and trials is indicated between parentheses."
expanded <- expand(attrition_bcn, lp, trial_type, is_valid_trial)

n_lp <- attrition_bcn |> 
	distinct(is_valid_participant, lp, session_id) |> 
	count(is_valid_participant, lp) |> 
	pivot_wider(names_from = is_valid_participant,
				values_from = n,
				names_repair = janitor::make_clean_names) |> 
	mutate(n_participants_true = true,
		   n_participants_false = false) |> 
	select(lp, true, false)

attrition_bcn |> 
	summarise(n_trials_true = sum(is_valid_trial),
			  n_trials_false = sum(!is_valid_trial),
			  .by = c(lp, trial_type)) |> 
	arrange(desc(lp)) |> 
	# across(n_trials, lst(mean, sd, sum), na.rm = TRUE),

	pivot_wider(id_cols = lp,
				names_from = c(trial_type),
				values_from = matches("trials"),
				names_repair = janitor::make_clean_names) |> 
	left_join(n_lp, by = join_by(lp)) |> 
	select(lp, true, false,
		   matches("n_trials")) |> 
	gt(rowname_col = "lp") |> 
	tab_spanner("Participants", matches("participant")) |> 
	tab_spanner("Trials", matches("trial")) |> 
	fmt_integer(everything()) |> 
	cols_merge(c(true, false),
			   pattern = "{1} ({2})") |> 
	cols_merge(c(n_trials_true_cognate,
				 n_trials_false_cognate),
				 pattern = "{1} ({2})") |> 
	cols_merge(c(n_trials_true_non_cognate,
				 n_trials_false_non_cognate),
				 pattern = "{1} ({2})") |> 
	cols_merge(c(n_trials_true_unrelated, 
				 n_trials_false_unrelated),
				 pattern = "{1} ({2})") |> 
	cols_label(true = "Participants", 
			   n_trials_true_cognate = "Cognate",
			   n_trials_true_non_cognate = "Non-cognate",
			   n_trials_true_unrelated = "Unrelated") |> 
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |>
	grand_summary_rows(fns = list(label = md("*N*")) ~ sum(.),
					   fmt = ~ fmt_integer(.))
```

@@

We collected data from `r n_participants_total` children (`r n_participants_sex$n[1]` female, `r n_participants_sex$n[2]` male, `r n_participants_sex$n[3]` sex not reported, age: *Mean* = `r round(ages$age_mean, 2)` months, *SD* = `r round(ages$age_sd, 2)`, *Range* = [`r round(ages$age_min, 2)`-`r round(ages$age_max, 2)`]), living in the Oxfordshire area (United Kingdom), tested at the Oxford BabyLab at the University of Oxford. Families were recruited from maternity rooms in private hospitals and social media, and contacted via phone when the child's age spanned between 20 and 32 months. From the `r n_participants_total` children that participated, `r n_participants_sessions$session_1` participated once, and `r n_participants_sessions$session_2` participated twice. Recurrent participants were tested with at least XXX months of difference. We gathered a total of `r n_sessions_total$n` testing sessions. All participants were being raised in exclusively British English monolingual homes.

@@

```{r dataset_oxf}
attrition_oxf <- filter(attrition, location=="Oxford")
n_trials <- nrow(attrition_oxf)
n_sessions <- n_distinct(attrition_oxf$session_id)
n_participants <- n_distinct(attrition_oxf$child_id)
n_trials_valid <- nrow(filter(attrition_oxf,
							  is_valid_participant,
							  is_valid_trial))
n_sessions_valid <- attrition_oxf |> 
	distinct(child_id, .keep_all = TRUE) |> 
	pull(is_valid_participant) |> 
	sum()

n_participants_valid <- n_distinct(
	attrition_oxf[attrition_oxf$is_valid_participant, ]$child_id)

n_exc_prime <- sum(!attrition_oxf$is_valid_gaze_prime)
n_exc_test <- sum(!attrition_oxf$is_valid_gaze_test)
n_exc_test_each <- sum(!attrition_oxf$is_valid_gaze_test_each)
n_exc_participants <- sum(!attrition_oxf |> 
	distinct(child_id, .keep_all = TRUE) |> 
	pull(is_valid_participant))

n_longitudinal <- attrition_oxf |>
	filter(is_valid_participant,
		   is_valid_trial) |>
	distinct(child_id, session_n) |> 
	count(child_id, name = "times") |> 
	count(times)

n_trials_condition <- attrition_oxf |> 
	filter(is_valid_participant) |> 
	count(trial_type, is_valid_trial) %>%
	split(janitor::make_clean_names(.$trial_type))
```

After trials that matched any of the aforementioned exclusion criteria from the dataset, we excluded participants who did not provide at least two valid trials in the *Cognate* condition, the *Non-cognate* condition, or the  *Unrelated* condition (*n* = `r n_exc_participants`). The final dataset included `r format(n_trials_valid, big.mark = ",")` trials from `r n_sessions_valid` testing sessions, generated by `r format(n_participants_valid, big.mark = ",")` distinct participants. Of those participants, `r n_longitudinal[1, 2]` provided data from one experimental session, `r n_longitudinal[2, 2]` provided data from two experimental sessions, and `r n_longitudinal[3, 2]` provided data from three experimental sessions. From the trials included in the final dataset dataset, `r n_trials_condition$unrelated_2$n` were *Unrelated* trials (`r n_trials_condition$unrelated$n` previously excluded), were `r n_trials_condition$non_cognate_2$n` *Related/Non-cognate* trials(`r n_trials_condition$non_cognate$n` previously excluded), and `r n_trials_condition$cognate_2$n` were *Related/Cognate* trials (`r n_trials_condition$cognate$n` previously excluded).
